Ecosystem size and complexity are extrinsic drivers of food chain length in branching networks

### 2022-10-26 rmd/manuscript.Rmd
Update manuscript.Rmd

We generated branching networks as connected habitat patches (**Figure 1A**). The geometric arrangement of networks was controlled by two parameters: ecosystem size, the number of patches ($N_{p}$); and ecosystem complexity, branching probability ($P_{b}$). Habitat patches were assigned to be a confluence (or upstream terminal patch) with probability $P_{b}$, or an in-branch patch with probability $1-P_{b}$. Branches (tributaries) are a series of connected patches between confluences (or upstream terminal patches). The number of patches within branch $q$, $n_{p,q}$, is a random variable drawn from a geometric distribution $n_{p,q} \sim Ge(P_{b})$ but conditional on $N_{p} = \sum_{q}^{N_{b}} n_{p,q}$ ($N_{b}$ is the number of branches in a network). Full details of the network generation process is described in @teruiEmergentDualScaling2021.

In our simulation, disturbance occurs simultaneously with probability $p_m$ but habitat patches vary in their vulnerability. We simulated the spatial distribution of disturbance intensity (i.e., proportional mortality in the event of stochastic regional disturbance) as follows. A disturbance value, $m_{h}$, was assigned randomly to upstream terminal patch $h$ as $logit~m_{h} \sim Normal(logit~\mu_m, \sigma_{h}^2)$, where $\mu_m$ is the average of disturbance intensity at headwaters in an ordinary scale. The disturbance values propagate downstream through a spatial autoregressive process as:

$$
logit~m_{down} = logit~m_{up}+ \epsilon
$$

where $m_{down}$ and $m_{up}$ are the disturbance values at down- and upstream patches, and the local environmental noise at the downstream patch $\epsilon$ was drawn from $Normal(0, \sigma_{l}^2)$. This autoregressive process resembles a natural disturbance regime, where flood or drought disturbance cascades downstream with some local variation along the channel [@nakamuraDisturbanceRegimesStream2000]. At confluence patches, the disturbance value takes a weighted mean based on the relative size of the two contributing branches, $Q$ (the number of upstream contributing patches + 1):

$$
logit~m_{down} = \omega~(logit~m_{1,up} +\epsilon_1)+(1-\omega)~(logit~m_{2,up} +\epsilon_2)
$$

$m_{1,up}$ and $m_{2,up}$ are the disturbance values of upstream contributing tributaries, $\epsilon_1$ and $\epsilon_2$ are independent local noises, and $\omega$ the relative size of one tributary over another defined as $\omega = \frac{Q_{1}}{Q_{1} + Q_{2}}$. Thus, the larger tributary has a greater influence on the downstream environment, as observed in natural river networks [@bendaNetworkDynamicsHypothesis2004]. Disturbance value distributions had medians close to $\mu_m$ with long tails (**Figure 1B**). The low-magnitude patches can be viewed as refugia, which supply individuals to impacted sites in subsequent time steps.

@@

$$
A_{i,x}(t+1) \sim Poisson(\bar{A}_{i,x}(t) + I_{i,x}(t) - E_{i,x}(t))
$$

@@

$$
\xi_{i,x}(t)=\frac{\sum_{y, y \ne x}E_{i,y}(t)exp(-\theta d_{xy})}{\sum_{x} \sum_{y, y \ne x}E_{i,y}(t)exp(-\theta d_{xy})}
$$

>>>>

We generated branching networks as connected habitat patches (**Figure 1A**). The geometric arrangement of networks was controlled by two parameters: ecosystem size, the number of patches ($N_{p}$); and ecosystem complexity, branching probability ($P_{b}$). Habitat patches were assigned to be a confluence (or upstream terminal patch) with probability $P_{b}$, or an in-branch patch with probability $1-P_{b}$. Branches (tributaries) are a series of connected patches between confluences (or upstream terminal patches). The number of patches within branch $q$, $n_{p,q}$, is a random variable drawn from a geometric distribution $n_{p,q} \sim \text{Ge}(P_{b})$ but conditional on $N_{p} = \sum_{q}^{N_{b}} n_{p,q}$ ($N_{b}$ is the number of branches in a network). Full details of the network generation process is described in @teruiEmergentDualScaling2021.

In our simulation, disturbance occurs simultaneously with probability $p_m$ but habitat patches vary in their vulnerability. We simulated the spatial distribution of disturbance intensity (i.e., proportional mortality in the event of stochastic regional disturbance) as follows. A disturbance value, $m_{h}$, was assigned randomly to upstream terminal patch $h$ as $\text{logit}~m_{h} \sim \text{Normal}(\text{logit}~\mu_m, \sigma_{h}^2)$, where $\mu_m$ is the average of disturbance intensity at headwaters in an ordinary scale. The disturbance values propagate downstream through a spatial autoregressive process as:

$$
\text{logit}~m_{down} = \text{logit}~m_{up}+ \varepsilon
$$

where $m_{down}$ and $m_{up}$ are the disturbance values at down- and upstream patches, and the local environmental noise at the downstream patch $\varepsilon$ was drawn from $\text{Normal}(0, \sigma_{l}^2)$. This autoregressive process resembles a natural disturbance regime, where flood or drought disturbance cascades downstream with some local variation along the channel [@nakamuraDisturbanceRegimesStream2000]. At confluence patches, the disturbance value takes a weighted mean based on the relative size of the two contributing branches, $Q$ (the number of upstream contributing patches + 1):

$$
\text{logit}~m_{down} = \omega~(\text{logit}~m_{1,up} +\varepsilon_1)+(1-\omega)~(\text{logit}~m_{2,up} +\varepsilon_2)
$$

$m_{1,up}$ and $m_{2,up}$ are the disturbance values of upstream contributing tributaries, $\varepsilon_1$ and $\varepsilon_2$ are independent local noises, and $\omega$ the relative size of one tributary over another defined as $\omega = \frac{Q_{1}}{Q_{1} + Q_{2}}$. Thus, the larger tributary has a greater influence on the downstream environment, as observed in natural river networks [@bendaNetworkDynamicsHypothesis2004]. Disturbance value distributions had medians close to $\mu_m$ with long tails (**Figure 1B**). The low-magnitude patches can be viewed as refugia, which supply individuals to impacted sites in subsequent time steps.

@@

$$
A_{i,x}(t+1) \sim \text{Poisson}(\bar{A}_{i,x}(t) + I_{i,x}(t) - E_{i,x}(t))
$$

@@

$$
\xi_{i,x}(t)=\frac{\sum_{y, y \ne x}E_{i,y}(t)\exp(-\theta d_{xy})}{\sum_{x} \sum_{y, y \ne x}\text{E}_{i,y}(t)\exp(-\theta d_{xy})}
$$
